---
title: "Miceplorer Vignette"
author: "Michael Gomez"
date: "2024-12-16"
output: pdf_document
---

First, let's download the package

```{r, eval = F}
devtools::install_github("mikeygomez/miceplorer")
```

Yay! It installed. Now, let's load it:

```{r, eval = F}
library(miceplorer)
library(readxl)
```

This vignette will use two approaches -> one using the clean_mouse_data to clean the data in one go and the other approach will clean the data by using each individual function.

## Approach 1

Let's load in and clean the data. If you want to do this, download the xlsx file from the data-raw file from github and replace path below. When console says to pick something from a table, the table is printed out as a knit-asis object.

```{r, eval = F}
path <- "~/miceplorer/data-raw/mousedata.xlsx"
cleaned_data <- clean_mouse_data(path)
```

Before, we continue it is important to generate a treatment_info dataframe which maps IDs to treatment groups.

```{r, eval = F}
treatment_info <- create_treatment_info(cleaned_data$birth)
```


Now, let's check for potential outliers in both outcomes and body weight.

```{r, eval = F}
weightoutliers <- identifyoutliers(data$weight, "Body Weight", treatment_info)
```
```{r, eval = F}
outcomeoutliers <- identifyoutliers(data$outcome, "Outcome", treatment_info)
```

Now, let's check for any significant weight loss.

```{r, eval = F}
weightcheck <- weightloss_check(cleaned_data$weight, threshold = 20)
```

Finally, we can generate a graph of both placebo and treatment looking at outcomes.

```{r, eval = F}
plots <- plot_indmeasurement(data$weight, "Body Weight", treatment_info)
print(plots$treatment)
print(plots$placebo)
```

## Approach 2

This approach will break up the clean_mouse_data() function into the separate functions that it calls to understand what's going on.

```{r, eval = F}
#maps sheets from the excel file to Birth, Body Weight, or Outcome
sheet_map <- excelsheetinfo(path)
```
Now, let's load in and clean the birth data set.

```{r, eval = F}
#reads in excel and call clean_birth_data function
birthdata <- readxl::read_excel(path, sheet = sheet_map[["Birth"]])
clean_birth_data <- clean_birth_data(birthdata)
```
Now let's load in the weight data. You can call clean_weight_data or break it up into its functions. First we'll use the clean_weight_data() function and then I'll break it up.

```{r, eval = F}
#reads in data and goes function by function
weightdata <- readxl::read_excel(path, sheet = sheet_map[["Body Weight"]])
clean_weight_data <- clean_weight_data(weightdata, clean_birth_data)
```

Broken up into its parts.

```{r, eval = F}
var <- "Body Weight"
data <- weightdata
#checks that column names are appropriate
columnnames <- check_columns(data, var)
data <- columnnames$data

#check that IDs are correct
IDcheck <- checkIDs(data, birthdata, var)
data <- IDcheck$data

#check that body weight values are numbers or NA
data <- checkquantvars(data, var)
```

Let's do a similar thing for outcome data.

```{r, eval = F}
outcomedata <- readxl::read_excel(path, sheet = sheet_map[["Outcome"]])
clean_outcome_data <- clean_outcome_data(outcomedata, clean_birth_data)
```

You can also break up the function like we did for birth weight, but since it would be redundant, I am omitting that.

Just like we did in approach 1, you can analyze the data further.

Congrats! You finished.
